### 2025-03-11 paper/paper.qmd

---
title: "`traveltime`: an R package to calculate travel time across a landscape from user-specified locations"
format:
  # md: default
  wordcount-html:
  # html:
    keep-md: true
    fig-height: 4
    fig-align: center
    fig-format: png
    dpi: 300
tags: 
  - R
  - geographic information systems
  - spatial analysis
authors:
  - name: Gerard E. Ryan
    orcid: 0000-0003-0183-7630
    corresponding: true
    affiliation: "1, 2"
  - name: Nicholas Tierney
    orcid: 0000-0003-1460-8722
    affiliation: "1, 3"
  - name: Nick Golding
    orcid: 0000-0001-8916-5570
    affiliation: "1, 4"
  - name: Daniel J. Weiss
    orcid: 0000-0002-6175-5648
    affiliation: "1, 3"
affiliations:
 - name: The Kids Research Institute Australia, Nedlands 6009 WA, Australia
   index: 1
 - name: Melbourne School of Population and Global Health, University of Melbourne, 3010, VIC, Australia
   index: 2
 - name: Curtin University, Bentley, WA, Australia
   index: 3
 - name: University of Western Australia, WA, Australia
   index: 4
date: In prep 2025
bibliography: paper.bib
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Summary

Understanding and mapping the time to travel among locations is useful for many activities from urban planning [@zahavi1974traveltime] to public health [@hulland2019travel; @weiss2020global] and myriad others [@nelson2019suite]. Here we present a software package --- `traveltime` --- written in and for the language R [@Rref]. `traveltime` enables a user to create a raster map of the travel time over an area of interest from a user-specified set of geographic coordinates. Applications include: estimating sampling bias [@dennis2000bias], allocating defibrillators [@tierney2018novel], setting health districts [@padgham2019introduction], or mapping access to vehicle chargers [@falchetta2021electric] and agricultural facilities [@zhao2023replanting].

The work-flow requires two key steps:

-   preparing a friction surface for the area of interest, and then
-   calculating travel time over that surface for the points of interest.

`traveltime` provides a spatial interface using object classes from the `terra` package [@terra]. Travel time is calculated as movement over a resistance 'friction surface' [@gdistance2017]. To provide easy access to the existing friction surfaces generated by @weiss2020global, `traveltime` internally uses `malariaAtlas` [@pfeffer2018malariaatlas] to download surfaces for the area of interest; though users can also supply any other friction surface.

`traveltime` is available from [R-Universe](https://idem-lab.r-universe.dev/traveltime) and [GitHub](https://github.com/idem-lab/traveltime), and documented at <https://idem-lab.github.io/traveltime/>. Although this article is intended to be the key reference for the `traveltime` package, we suggest citations of the package are accompanied by citing the underlying methodological work as well [@weiss2018global; @weiss2020global].

# Statement of need

Global maps of travel time to cities [@weiss2018global; @nelson2019suite] and health care [@hulland2019travel; @weiss2020global] have generated much interest and use[^1], and the city data set of @nelson2019suite is available to R users through the widely-used `geodata` package [@geodata]. Here we extend that work to enable travel time calculations from any arbitrary set of locations and friction surface.

[^1]: Collectively >1600 citations per Google Scholar at the 28th of January 2025.

A gaggle of R packages provide superficially similar though fundamentally different functionality via the [TravelTime.com](https://www.TravelTime.com) API [@traveltimeapi; @traveltimeR; @rtraveltime; @traveltime_gh]. Their 'Isochron' polygons --- areas reachable within a given time from a given location --- are most comparable to what `traveltime::calculate_travel_time()` calculates. However, each isochron is a single polygon calculated is for a single point location and specified maximum travel time, rather than a raster of temporal gradation across a landscape, jointly for an arbitrary number of points, as in `traveltime`.

With `traveltime`, we provide free and open source software to estimate travel time from any number of user-supplied locations, across a complete area of interest, and with convenient access to motorised transport or walking friction surfaces with global coverage.

# Example: walking from public transport in Singapore

Here we provide a very brief example to calculate and map the walking travel time from rail transport stations across Singapore. [Complete code for this example is available as a vignettte in package documentation](https://idem-lab.github.io/traveltime/articles/singapore.html).

## Prepare data and friction surface

We need two items of data:

-   our area of interest --- Singapore, and
-   our points to calculate travel time from --- Singapore's Mass Rapid Transit (MRT) and Light Rail Transit (LRT) stations.

We download `singapore_shapefile`, a polygon boundary of Singapore from the GADM [@gadm] database using the `geodata` package [@geodata] as our area of interest. Our points are the `traveltime::stations` data, containing coordinates of all LRT and MRT station exits in Singapore [@singdata]. We then download a walking travel friction surface for Singapore:

```{r}
#| echo: false
#| output: false
#| label: get-basemap
library(terra)
library(geodata)

singapore_shapefile <- gadm(
  country = "Singapore",
  level = 0,
  path = tempdir(),
  resolution = 2
)

singapore_shapefile
```

```{r}
#| echo: false
#| output: false
#| label: show-stations
library(traveltime)
head(stations)
```

```{r}
#| label: download-friction-surface
#| echo: true
#| output: false
friction_singapore <- traveltime::get_friction_surface(
    surface = "walk2020",
    extent = singapore_shapefile
  )|> 
  terra::mask(singapore_shapefile)
```

(Alternatively, we could use `surface = "motor2020"` for motorised travel. We're also only interested in walking *on land* so we `mask` out areas outside of the land boundary in `singapore_shapefile`)

```{r }
#| echo: false
#| output: false
#| label: show-friction-surface-raster
friction_singapore
```

We plot these data below. `traveltime` takes resistance values of friction [@gdistance2017], so higher values of friction indicate more time travelling across a given cell.

```{r}
#| label: fig-data
#| echo: false
#| fig-cap: "Friction surface raster of Singapore, showing Singapore boundary
#|   in grey, and station locations as grey points."
library(tidyterra)
library(ggplot2)

ggplot() +
  geom_spatraster(
    data = friction_singapore
  ) +
  geom_spatvector(
    data = singapore_shapefile,
    fill = "transparent",
    col = "grey50"
  ) +
  geom_point(
    data = stations,
    aes(
      x = x,
      y = y
    ),
    col = "grey60",
    size = 0.5
  ) +
  scale_fill_viridis_c(
    option = "A",
    na.value = "transparent",
    direction = -1
  ) +
  labs(
    fill = "Resistance",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_minimal()
```

## Calculate and plot the travel time

With all the data collected, the function `calculate_travel_time()` takes the friction surface `friction_singapore` and the points of interest in `stations`, and returns a `SpatRaster` of walking time in minutes to each cell from the nearest station:

```{r}
#| label: calculate-result
trave_time_singapore <- calculate_travel_time(
  friction_surface = friction_singapore,
  points = stations
)

trave_time_singapore
```

Resulting travel times are Figure \ref{fig-result}[^2].

```{r}
#| label: fig-result
#| echo: false
#| fig-cap: "Map of walking travel time in Singapore, in minutes from 
#|   nearest MRT or LRT station."
ggplot() +
  geom_spatraster(
    data = trave_time_singapore
  ) +
  scale_fill_viridis_c(
    option = "A",
    direction = -1,
    na.value = "transparent"
  ) +
  theme_minimal() +
  labs(fill = "Minutes") +
  geom_spatvector(
    data = singapore_shapefile,
    fill = "transparent",
    col = "grey20"
  )
```

[^2]: In Figure \ref{fig-data}, the islands to the south and north-east are filled, but unconnected with the mainland. These cells appear absent in Figure \ref{fig-result}; because they are not connected to any cells with a station, the calculated travel time is infinite, and so are not plotted in Figure \ref{fig-result}.

# Opportunities for future development

The `traveltime` package is immediately suitable to a range of application. Nonetheless, we see opportunities to build the package utility through:

  - capability to better distribute a wider range friction surfaces, and
  - additional methods to efficiently compute results over large spatial extents.

# Acknowledgements

This work was supported, in whole or in part, by the Bill & Melinda Gates Foundation [INV-021972]. The conclusions and opinions expressed in this work are those of the authors alone and shall not be attributed to the Foundation. Under the grant conditions of the Foundation, a Creative Commons Attribution 4.0 License has already been assigned to the Author Accepted Manuscript version that might arise from this submission. Please note works submitted as a preprint have not undergone a peer review process.

# References